---
title: "Song Recommendation Algorithms Using K-Means, KNN, and Spotify Echo Nest"
author: "Alina Hagen, Daniel Tafmizi, Alexander Crowell, Anthony Venson"
date: "`r Sys.Date()`"
output: html_document
runtime: shiny
---

## Data Munging 

``` {r, "DATA COMBINATION"}
library(readr)
library(tidyverse)
library(dbplyr)
songs <- read_csv("enchonest_2.csv")
songs_2 <- as.data.frame(read_csv("hit_songs_09_19.csv"))
songs_3 <- read_csv("echonest_3.csv")

deduped_songs <- songs %>%
  arrange(desc(popularity)) %>% 
  distinct(artists, track_name, .keep_all = TRUE) %>%
  select(artists, track_name, duration_ms, popularity, danceability, energy, key, loudness, mode, speechiness, acousticness,instrumentalness, liveness, valence, tempo)

songs_2 <- songs_2 %>%
  select(artist, song, duration_ms, popularity, danceability, energy, key, loudness, mode, speechiness, acousticness,instrumentalness, liveness, valence, tempo)

deduped_songs2 <- songs_3 %>%
  arrange(desc(popularity)) %>% 
  mutate(
    artist = str_remove_all(artists, "\\[|\\]|'"),   
    artist = str_trim(artist),
    track_genre = NA) %>%
  distinct(artists, name, .keep_all = TRUE) %>%
  select(artist, name, duration_ms, popularity, danceability, energy, key, loudness, mode, speechiness, acousticness,instrumentalness, liveness, valence, tempo)

colnames(deduped_songs2) <- c("artists", "track_name", "duration_ms", "popularity", "danceability", "energy", "key", "loudness", "mode", "speechiness", "acousticness","instrumentalness", "liveness", "valence", "tempo")

colnames(songs_2) <- c("artists", "track_name", "duration_ms", "popularity", "danceability", "energy", "key", "loudness", "mode", "speechiness", "acousticness","instrumentalness", "liveness", "valence", "tempo")

df_new_only <- anti_join(deduped_songs2, songs_2)
df_combined <- bind_rows(songs_2, df_new_only)
df_new_only2 <- anti_join(deduped_songs, df_combined)
df_combined <- bind_rows(df_combined, df_new_only2)


df_combined <- df_combined %>% arrange(desc(popularity)) %>% filter(popularity >50) %>% distinct(artists, track_name, .keep_all = TRUE) %>% distinct(duration_ms, track_name, .keep_all = TRUE) 

write.csv(df_combined, "df_combined.csv")

```

## Load and Prepare Data
### Look at Data Correlation

``` {r, "DATA LOADING"}
library(readr)
library(tidyverse)
library(dbplyr)

df_combined <- read_csv("df_combined.csv")

# extract song names
song_name <- df_combined[,(2:3)]
# Set all as numeric
songs <- df_combined[c(-1:-3)]
songs[] <- lapply(songs, function(x) as.numeric(as.character(x)))
```

``` {r, "CORRELATION"}
library(corrplot)

corrmatrix <- cor(songs)
corrplot(corrmatrix, method = 'number')

```

``` {r, "Search Test"} 
which(df_combined$track_name == "Another Love")
```

## MODEL 1: SIMPLE KNN ANALYSIS

``` {r, "MODEL 1"}
if (!require(FNN)) {
  install.packages("FNN") 
  library(FNN)
}

library(FNN)

# 1. Scale data
songs_scaled <- scale(songs)

# 2. Pick point
query_point <- songs_scaled[843, , drop = FALSE]

# 3. Find 5 nearest neighbors to point (Report will include the song+5 nearest neighbors so k = 6)
knn_result <- get.knnx(data = songs_scaled, query = query_point, k = 6)

# 4. Inspect the neighbor indices and distances
neighbor_indices <- knn_result$nn.index[]
neighbor_distances <- knn_result$nn.dist[]

# 5. Show actual neighbor rows
neighbors_data <- song_name[neighbor_indices[1, ], ]

# Print everything
cat("Nearest neighbors (indices):\n")
print(neighbor_indices)

cat("\nEuclidian Distances:\n")
print(neighbor_distances)

cat("\nNeighbor data:\n")
print(neighbors_data)
```

## K CLUSTERING DETERMINATION 
``` {r, "K-CLUSTERING"}
library(FNN)
library(cluster)
library(tidyverse)
library(reshape2)
set.seed(42)

# 1. Scale data
songs_scaled <- scale(songs)

sampled_df <- songs %>% slice_sample(prop = 0.10)
sample <- scale(sampled_df)

# Two Method K- Cluster Verification 
############################################################
## OPTIMAL K ANALYSIS VIA DERIVIATION AND ELBOW DETECTION ##
############################################################

# Compute WSS valus
wss <- map_dbl(1:15,  function(k){
  model <- kmeans(x = sample, centers = k, nstart = 50, iter.max = 100)
  model$tot.withinss
})

# Emulate Deriviation Logic with a Descrete function using diff()
# Compute first and second differences
first_deriv <- diff(wss)
second_deriv <- diff(first_deriv)

# Find where the second derivative (Change in first derivative) is below 10%
elbow_k <- which.max(abs(second_deriv)) + 1
cat("Optimal k (via second derivative):", elbow_k, "\n")

# Elbow Plot
plot(1:15, wss, type = "b", pch = 19, 
     main = "WSS vs. K with Elbow with Automatic Detection via Deriviation",
     xlab = "Number of Clusters (k)", ylab = "WSS")
abline(v = elbow_k, col = "red", lty = 2)
text(elbow_k, wss[elbow_k], labels = paste0("k=", elbow_k), pos = 4, col = "red")

##############################################
## OPTIMAL K ANALYSIS VIA SILHOUETTE METHOD ##
##############################################

silhouette_scores <- c()

for (k in 2:12) {
  km <- kmeans(sample, centers = k, nstart = 50)
  sil <- silhouette(km$cluster, dist(sample))  # FIXED HERE
  avg_sil_width <- mean(sil[, 3])   # average silhouette width
  silhouette_scores <- c(silhouette_scores, avg_sil_width)
}

silhouette_k <- which.max(silhouette_scores) + 1 
cat("Optimal k (via silhoutte analysis):", silhouette_k, "\n")

plot(2:12, silhouette_scores, type = "b",
     xlab = "Number of clusters (k)",
     ylab = "Average silhouette width",
     main = "Choosing k with Silhouette Scores")
points(silhouette_k, silhouette_scores[silhouette_k - 1], col = "red", pch = 19, cex = 1.5)
text(silhouette_k, silhouette_scores[silhouette_k - 1], labels = paste("Best k =", silhouette_k),
     pos = 3, col = "red", cex = 0.9)

########################################
## ELBOW METHOD == SILHOUETTE METHOD? ##
########################################
if (elbow_k == silhouette_k) {
  k = silhouette_k
  cat("Elbow method and Silhouette method agree: optimal k =", k, ".\n")
} else {
  k <- silhouette_k
  cat("Warning: Elbow and Silhouette suggest different k values.\n")
}

#############################################
## K MEANS CLUSTERING WITH OPTIMAL K VALUE ##
#############################################

kmeans_result <- kmeans(songs_scaled, centers = k)

# Reshape Data to long formt via reshape2
centers_df <- as.data.frame(kmeans_result$centers)
centers_df$cluster <- rownames(centers_df)
centers_long <- pivot_longer(centers_df, -cluster, names_to = "feature", values_to = "value")

# Visualize Cluster Difference Analysis
ggplot(centers_long, aes(x = feature, y = value, group = cluster, color = cluster)) +
  geom_line() +
  geom_point() +
  theme_minimal() +
  coord_flip() +
  labs(title = "Cluster Centroid Feature Values", y = "Scaled Value")

```

``` {r, "CLUSTERING PLOT"}
library(GGally)
# Convert to data frame and add cluster
plot_df <- as.data.frame(songs_scaled) %>%
  mutate(cluster = factor(kmeans_result$cluster))

# Select the features you want to include
metrics <- c("popularity", "danceability", "energy", "loudness", "valence")

# Generate the pairplot matrix
ggpairs(
  plot_df,
  columns = which(colnames(plot_df) %in% metrics),
  aes(color = cluster, alpha = 0.6),
  upper = list(continuous = wrap("points", size = 0.7)),
  lower = list(continuous = wrap("points", size = 0.7)),
  diag = list(continuous = wrap("densityDiag"))
) +
  theme_minimal()

```


## MODEL 2: K-Clustering (K determined via Elbow and Silhouette) then cluster focused KNN

``` {r, "MODEL 2"}
##################################################
## K NEAREST NEIGHBOR WITHIN DETERMINED CLUSTER ##
##################################################

# Pick a Point
query_index <- 843

# Identify query's cluster and all songs in that cluster
query_cluster <- kmeans_result$cluster[query_index]
cluster_indices <- which(kmeans_result$cluster == query_cluster)

# Subset cluster data
cluster_songs <- songs_scaled[cluster_indices, , drop = FALSE]
cluster_song_names <- song_name[cluster_indices, ]

# Get the query song within the cluster
relative_query_index <- match(query_index, cluster_indices)
query_cluster_point <- cluster_songs[relative_query_index, , drop = FALSE]

# 5 nearest neighbors (Report will include the song+5 nearest neighbors so k = 6)
knn_result <- get.knnx(data = cluster_songs, query = query_cluster_point, k = 6)

# Neighbor Results 
neighbor_indices <- knn_result$nn.index[1, ]
neighbor_distances <- knn_result$nn.dist[1, ]
neighbors_data <- cluster_song_names[neighbor_indices, ]


####################
## RESULTS REPORT ##
####################

cat("Nearest neighbors (indices):\n")
print(neighbor_indices)

cat("\nEuclidian Distances:\n")
print(neighbor_distances)

cat("\nNeighbor data:\n")
print(neighbors_data)
```

## MODEL 3: K-Clustering (K determined via Elbow and Silhouette), PCA Dimensionality reduction,  then cluster focused KNN with PCA variables

``` {r, "PCA"}
##################################
## PCA DIMENSIONALITY REDUCTION ##
##################################
pca_result <- prcomp(songs_scaled, scale. = FALSE)  # Already scaled, so scale.=FALSE

explained_variance <- (pca_result$sdev)^2 / sum(pca_result$sdev^2)
cumulative_variance <- cumsum(explained_variance)

# While I could not find a 100% accepted rule of thumb for acceptable variability thresholds,
# Shmueli et al. (2017) appears to refer back to roughly 90% variance often in Ch. 4 (Dimensionality   
# Reduction, page 108), So I am going to use 90% as the variance threshold
plot(cumulative_variance, type = "b",
     xlab = "Number of Principal Components",
     ylab = "Cumulative Explained Variance",
     main = "Explained Variance by Principal Components")
abline(h = 0.90, col = "red", lty = 2)  

num_pcs <- which(cumulative_variance >= 0.90)[1]  
songs_pca <- pca_result$x[, 1:num_pcs]  # Top 5 principal components

```

``` {r, "MODEL 3"}
#####################################################################
## K NEAREST NEIGHBOR WITHIN DETERMINED CLUSTER WITH PCA REDUCTION ##
#####################################################################

# Pick a Point
query_index <- 843

# Identify query's cluster and all songs in that cluster
query_cluster <- kmeans_result$cluster[query_index]
cluster_indices <- which(kmeans_result$cluster == query_cluster)

# Subset cluster data
cluster_songs <- songs_pca[cluster_indices, , drop = FALSE]
cluster_song_names <- song_name[cluster_indices, ]

# Get the query song within the cluster
relative_query_index <- match(query_index, cluster_indices)
query_cluster_point <- cluster_songs[relative_query_index, , drop = FALSE]

# 5 nearest neighbors (Report will include the song+5 nearest neighbors so k = 6)
knn_result <- get.knnx(data = cluster_songs, query = query_cluster_point, k = 6)

# Neighbor Results 
neighbor_indices <- knn_result$nn.index[1, ]
neighbor_distances <- knn_result$nn.dist[1, ]
neighbors_data <- cluster_song_names[neighbor_indices, ]


####################
## RESULTS REPORT ##
####################

cat("Nearest neighbors (indices):\n")
print(neighbor_indices)

cat("\nEuclidian Distances:\n")
print(neighbor_distances)

cat("\nNeighbor data:\n")
print(neighbors_data)
```

## SPOTIFY API LOOKUP ##

``` {r, "SPOTIFY API KEYS", include = FALSE}
library(spotifyr)

# Authenticate (client credentials is fine)
Sys.setenv(SPOTIFY_CLIENT_ID = '5b83c7fe870042b492a35d75da13680a')
Sys.setenv(SPOTIFY_CLIENT_SECRET = '4c27d19dfdec4b0085838cd0f87fd867')
Sys.setenv(SPOTIFY_REDIRECT_URI = "http://127.0.0.1:1410/")
```

``` {r, "LOOK UP LIST"}
library(stringr)

lookup_list <- neighbors_data %>%
  mutate(combined = str_c(artists, track_name, sep = " "))

liked <- lookup_list[1,]
lookup_list <- lookup_list[-1,]

```

``` {r, results='asis'}
access_token <- get_spotify_access_token()

lookup_list$spotify_url <- NA

for (i in seq_along(lookup_list$combined)) {
  query <- lookup_list$combined[i]

  result <- tryCatch({
    search_spotify(query, type = "track", limit = 1, market = "US")
  }, error = function(e) NULL)

  if (!is.null(result) && nrow(result) > 0) {
    lookup_list$spotify_url[i] <- result$external_urls.spotify[1]
  }
}

recommendations <- paste0(
  'Try "', lookup_list$track_name, '" by ', lookup_list$artists,
  ' [Spotify link](', lookup_list$spotify_url, ')', "\n \n \n"
)

name <- ""  
name <- paste0(name, liked[1,2])
artist <- ""  
artist <- paste0(artist, liked[1,1])

cat("Since you liked", name, "by", artist, "\n\n",
  recommendations)

```





## ANNOTATION CODE ##
## BATCH QUERY FUNCTIONS ##

```{r, "MODEL 1 Function"}
library(FNN)
library(dplyr)

batch_knn_standard <- function(query_indices, songs_scaled, song_name, k) {
  all_results <- list()

  for (query_index in query_indices) {
    query_point <- songs_scaled[query_index, , drop = FALSE]
    knn_result <- get.knnx(data = songs_scaled, query = query_point, k = k + 1)  
    
    neighbor_indices <- knn_result$nn.index[1, ]
    neighbor_distances <- knn_result$nn.dist[1, ]

    # Remove first Neighhor (Self)
    neighbor_indices <- neighbor_indices[-1]
    neighbor_distances <- neighbor_distances[-1]
    
    # Extract Song Names
    neighbors_data <- song_name[neighbor_indices, , drop = FALSE]

    # 5. Save neighbors with query_id
    result_df <- data.frame(
      query_id = query_index,
      neighbor_text = apply(neighbors_data, 1, paste, collapse = " - "),
      stringsAsFactors = FALSE
    )

    all_results[[as.character(query_index)]] <- result_df
  }

  final_df <- bind_rows(all_results)
  return(final_df)
}

```

```{r, "MODEL 2 and 3 Function"}
library(FNN)
library(dplyr)

batch_clustered_knn <- function(query_indices, kmeans_result, data, song_name, k = 6) {
  all_results <- list()

  for (query_index in query_indices) {
    # Identify cluster of the query point
    query_cluster <- kmeans_result$cluster[query_index]
    cluster_indices <- which(kmeans_result$cluster == query_cluster)

    # Get cluster-specific data
    cluster_songs <- data[cluster_indices, , drop = FALSE]
    cluster_song_names <- song_name[cluster_indices, ]

    # Locate query within the cluster
    relative_query_index <- match(query_index, cluster_indices)
    query_point <- cluster_songs[relative_query_index, , drop = FALSE]

    # Run KNN in the cluster
    knn_result <- get.knnx(data = cluster_songs, query = query_point, k = k + 1)

    neighbor_indices <- knn_result$nn.index[1, ]
    neighbor_indices <- neighbor_indices[neighbor_indices != relative_query_index]  # remove self
    neighbors_data <- cluster_song_names[neighbor_indices, , drop = FALSE]

    # Build results with query ID
    result_df <- data.frame(
      query_id = query_index,
      neighbor_text = apply(neighbors_data, 1, paste, collapse = " - "),
      stringsAsFactors = FALSE
    )

    all_results[[as.character(query_index)]] <- result_df
  }

  final_df <- bind_rows(all_results)
  return(final_df)
}

```

```{r, "SPOTIFY API CALL"}
library(dplyr)
library(tidyr)
library(stringr)
library(spotifyr)

format_neighbors_with_spotify_urls <- function(neighbors_df, k = 5) {
  access_token <- get_spotify_access_token()

  # Step 1: Extract artist and track name
  split_df <- neighbors_df %>%
    mutate(rank = row_number(),
           artist = str_trim(str_extract(neighbor_text, "^[^–-]+")),
           track_name = str_trim(str_remove(neighbor_text, "^[^–-]+[-–]")))

  # Step 2: Lookup Spotify URLs
  get_spotify_url <- function(artist, track) {
    query <- paste(artist, track)
    result <- tryCatch({
      search_spotify(query, type = "track", limit = 1, market = "US")
    }, error = function(e) NULL)

    if (!is.null(result) && nrow(result) > 0) {
      return(result$external_urls.spotify[1])
    } else {
      return(NA)
    }
  }

  split_df$spotify_url <- mapply(get_spotify_url, split_df$artist, split_df$track_name)

  # Step 3: Rank neighbors per query_id
  split_df <- split_df %>%
    group_by(query_id) %>%
    mutate(rank = row_number()) %>%
    filter(rank <= k) %>%
    ungroup()

  # Step 4: Pivot to wide format
  formatted <- split_df %>%
    mutate(
      rec_col = paste0("Model Rec ", rank),
      url_col = paste0("Song URL ", rank),
      annotation_col = paste0("Annotation ", rank)
    ) %>%
    select(query_id, rank, neighbor_text, spotify_url) %>%
    pivot_wider(
      names_from = rank,
      values_from = c(neighbor_text, spotify_url),
      names_glue = "{.value} {rank}"
    )

  # Step 5: Add empty annotation columns
  for (i in 1:k) {
    formatted[[paste0("Annotation ", i)]] <- ""
  }

  # Step 6: Reorder to match Excel layout
  formatted <- formatted %>%
    relocate(query_id,
      starts_with("neighbor_text 1"), starts_with("spotify_url 1"), starts_with("Annotation 1"),
      starts_with("neighbor_text 2"), starts_with("spotify_url 2"), starts_with("Annotation 2"),
      starts_with("neighbor_text 3"), starts_with("spotify_url 3"), starts_with("Annotation 3"),
      starts_with("neighbor_text 4"), starts_with("spotify_url 4"), starts_with("Annotation 4"),
      starts_with("neighbor_text 5"), starts_with("spotify_url 5"), starts_with("Annotation 5")
    )

  # Optional: Rename columns for final polish
  names(formatted) <- names(formatted) %>%
    str_replace("neighbor_text", "Model Rec") %>%
    str_replace("spotify_url", "Song URL")

  return(formatted)
}

```

``` {r}
# Query songs by index (e.g., row numbers in your scaled data)
query_set <- c(34,5253,28496,5851,9422,11895,9135,46249,28518,3176) # ALI
# query_set <- c(13,16,324,2055,1384,29,1235,1433,108,370) # DANIEL
# query_set <- c(454,162,194,770,1419,9121,9142,5126,26344,5930) #ALEX
# query_set <- c(57,58,99,129,355,5103,9608,12844,19342) #ANTHONY


# Run batch standard KNN
neighbors_df_s <- batch_knn_standard(
  query_indices = query_set,
  songs_scaled = songs_scaled,
  song_name = song_name,
  k = 6
)

# Run batch clustered KNN
neighbors_df_c <- batch_clustered_knn(
  query_indices = query_set,
  kmeans_result = kmeans_result,
  data = songs_scaled,
  song_name = song_name,
  k = 6
)

# Run batch clustered KNN + PCA
neighbors_df_pca <- batch_clustered_knn(
  query_indices = query_set,
  kmeans_result = kmeans_result,
  data = songs_pca,
  song_name = song_name,
  k = 6
)

View(neighbors_df_c)

```

```{r}
model1 <- format_neighbors_with_spotify_urls(neighbors_df_s, k = 5)
model2 <- format_neighbors_with_spotify_urls(neighbors_df_c, k = 5)
model3 <- format_neighbors_with_spotify_urls(neighbors_df_pca, k = 5)

#write.csv(model1, "x1.csv")
#write.csv(model2, "x2.csv")
#write.csv(model3, "x3.csv")

``` 


## EVALUATION RESULTS ##
|      | Model 1 Songs Liked | Model 1 Songs Disliked | Model 1 Final Percent | Model 2 Songs Liked | Model 2 Songs Disliked | Model 2 Final Percent | Model 3 Songs Liked | Model 3 Songs Disliked | Model 3 Final Percent |
|------|---------------------|-------------------------|------------------------|---------------------|-------------------------|------------------------|---------------------|-------------------------|------------------------|
| Alina   | 26                  | 24                      | 0.52                   | 28                  | 22                      | 0.56                   | 33                  | 17                      | 0.66                   |
| Daniel   | 42                  | 8                       | 0.84                   | 42                  | 8                       | 0.84                   | 37                  | 13                      | 0.74                   |
| Alex   | 23                  | 27                      | 0.46                   | 25                  | 25                      | 0.50                   | 23                  | 27                      | 0.46                   |
| Anthony   | 27                  | 23                      | 0.54                   | 28                  | 22                      | 0.56                   | 25                  | 25                      | 0.50                   |

``` {r}
library(ggplot2)


music_results <- read.csv("Book1.csv")

colnames(music_results) <- c('Name', 'Model1', 'Model2', 'Model3')

finalResults <- c(0.59, 0.615, 0.59)
finalModels <- c("Model 1", "Model 2", "Model 3")
final_df <- data.frame(finalModels, finalResults)

#Individual Plots

ggplot(music_results, aes(x = Name, y = Model1)) +
  geom_col(data = music_results, aes(fill = Model1)) +
  theme(legend.position = "none") +
  labs(x = "Name", y = "Model Accuracy (%)", title = "Model 1 Prediction Accuracy") +
  coord_flip() 

ggplot(music_results, aes(x = Name, y = Model2)) +
  geom_col(data = music_results, aes(fill = Model2)) +
  theme(legend.position = "none") +
  labs(x = "Name", y = "Model Accuracy (%)", title = "Model 2 Prediction Accuracy") +
  coord_flip()

ggplot(music_results, aes(x = Name, y = Model3)) +
  geom_col(data = music_results, aes(fill = Model3)) +
  theme(legend.position = "none") +
  labs(x = "Name", y = "Model Accuracy (%)", title = "Model 3 Prediction Accuracy") +
  coord_flip()

ggplot(final_df, aes(x = finalModels, y = finalResults)) +
  geom_col(data = final_df, aes(fill = finalResults)) +
  theme(legend.position = "none") +
  labs(x = "Model", y = "Model Accuracy (%)", title = "Final Prediction Accuracy")


#Faceted Plot

faceted_df <- read.csv("Book2.csv")
colnames(faceted_df) <- c("Name", "Model", "Accuracy")

ggplot(faceted_df, aes(x = Model, y = Accuracy)) +
  geom_bar(stat = "identity", position = "dodge", aes(fill = Name)) +
  labs(x = "Model", y = "Accuracy (%)", title = "Predictive Model Success Rate") +
  theme_classic()

ggplot(faceted_df, aes(x = Model, y = Accuracy)) +
  geom_bar(stat = "identity", aes(fill = Name)) +
  labs(x = "Model", y = "Combined Accuracy", title = "Total Model Success Rate") +
  theme_classic()

```

## Shiny App Code 

```{r, "SHINY APP"}

library(shiny)
library(shinythemes)
library(FNN)
library(stringdist)
library(httr)
library(base64enc)

# Load data
clust1 <- read.csv("cluster1.csv")
clust2 <- read.csv("cluster2.csv")

songs1 <- read.csv("song_name_cluster1.csv")
songs2 <- read.csv("song_name_cluster2.csv")

# Creates lookup table
lookup <- data.frame(
  song_name = c(songs1$full_name, songs2$full_name),
  cluster_id = c(rep(1, nrow(songs1)), rep(2, nrow(songs2))),
  stringsAsFactors = FALSE
)

# Fuzzy matching function
find_closest_song <- function(input_song_name, lookup_table) {
  distances <- stringdist::stringdist(tolower(input_song_name), tolower(lookup_table$song_name), method = "jw")
  sorted_indices <- order(distances)
  top_indices <- sorted_indices[1:6]
  top_distances <- distances[top_indices]
  top_names <- lookup_table$song_name[top_indices]
  top_clusters <- lookup_table$cluster_id[top_indices]
  
  return(data.frame(song_name = top_names, distance = top_distances, cluster_id = top_clusters))
}

# Function to get a new access token
get_spotify_access_token <- function(client_id, client_secret) {
  credentials <- paste(client_id, client_secret, sep = ":")
  encoded_credentials <- base64encode(charToRaw(credentials))
  
  response <- POST(
    url = "https://accounts.spotify.com/api/token",
    add_headers(Authorization = paste("Basic", encoded_credentials)),
    body = list(grant_type = "client_credentials"),
    encode = "form"
  )
  
  token_info <- content(response)
  return(token_info$access_token)
}

# UI
ui <- fluidPage(
  theme = shinytheme("darkly"),  # Dark theme
  
  tags$style(HTML("
    .centered-input {
      width: 100%;
      display: flex;
      justify-content: center;
    }
    .centered-input input {
      text-align: center;
      width: 100%;
    }
  ")),
  
  tags$div(
    style = "text-align: center; margin-top: 30px;",
    tags$h1(
      "Song Finder",
      style = "color: white;"
    )
  ),
  
  fluidRow(
    column(
      width = 12,
      align = "center",
      div(
        style = "width: 50%; margin: auto;",
        div(
          class = "centered-input",
          textInput("song_name", 
                    label = "Enter Song Name: ", 
                    placeholder = "Song Name - Artist")
        ),
        actionButton("find_btn", "Find Song"),
        br(), br(),
        uiOutput("advanced_ui")
      ),
      br(),
      div(
        style = "width: 50%; margin: auto;",
        textOutput("current_song_name"), 
        tableOutput("neighbors")
      ),
      div(
        style = "width: 50%; margin: auto; margin-top: 20px;",
        uiOutput("alternative_songs_ui")
      ),
      br(), br(),
      div(
        style = "width: 100%; text-align: center; margin-top: 40px; color: grey;",
        tags$h5("Database is not exhaustive due to new Spotify API restrictions. Limited songs after 2022.")
      )
    )
  )
)

# Server
server <- function(input, output, session) {
  
  token_reactive <- reactiveVal()
  top_matches <- reactiveVal()
  best_match <- reactiveVal()
  selection_made <- reactiveVal(FALSE)
  knn_done <- reactiveVal(FALSE)
  
  #fill with api keys
  client_id <- ''
  client_secret <- ''
  
  access_token <- get_spotify_access_token(client_id, client_secret)
  token_reactive(access_token)
  
  # Spotify link search function
  get_spotify_link <- function(song_title_artist) {
    query <- URLencode(song_title_artist)
    url <- paste0("https://api.spotify.com/v1/search?q=", query, "&type=track&limit=1&market=US")
    
    response <- GET(
      url,
      add_headers(Authorization = paste("Bearer", token_reactive()))
    )
    
    result <- content(response)
    
    if (!is.null(result$tracks$items) && length(result$tracks$items) > 0) {
      return(result$tracks$items[[1]]$external_urls$spotify)
    } else {
      return(NA)
    }
  }
  
  #function to run knn
  perform_knn <- function(selected_match) {
    matched_song_name <- selected_match$song_name
    query_cluster_id <- selected_match$cluster_id
    
    if (query_cluster_id == 1) {
      cluster_data <- clust1
      cluster_song_names <- songs1
      relative_query_index <- which(songs1$full_name == matched_song_name)
    } else {
      cluster_data <- clust2
      cluster_song_names <- songs2
      relative_query_index <- which(songs2$full_name == matched_song_name)
    }
    
    knn_result <- get.knnx(
      data = cluster_data,
      query = cluster_data[relative_query_index, , drop = FALSE],
      k = 6
    )
    
    neighbor_indices <- knn_result$nn.index[1, 2:6]
    neighbor_distances <- knn_result$nn.dist[1, 2:6]
    neighbors_data <- cluster_song_names[neighbor_indices, ]
    
    return(list(
      indices = neighbor_indices,
      distances = neighbor_distances,
      song_names = neighbors_data$full_name
    ))
  }
  # runs the knn function when find button is clicked
  observeEvent(input$find_btn, {
    input_song_name <- input$song_name
    matches <- find_closest_song(input_song_name, lookup)
    
    top_matches(matches)
    best_match(matches[1, ])
    
    selection_made(FALSE)
    knn_done(FALSE)
    
    output$current_song_name <- renderText({
      paste("Showing Recommendations for:", best_match()$song_name)
    })
    
    output$alternative_songs_ui <- renderUI({
      selectInput(
        "alternative_song", 
        "Not the right song? Check Below:", 
        choices = c("Alternative Matches" = "", matches$song_name[2:6]),
        selected = NULL
      )
    })
    
    knn_results <- perform_knn(best_match())
    knn_done(TRUE)
    
    spotify_links <- sapply(knn_results$song_names, get_spotify_link)
    
    output$neighbors <- renderTable({
      data.frame(
        Order = 1:5,
        Song_Name = knn_results$song_names,
        URLs = sprintf('<a href="%s" target="_blank">Link</a>', spotify_links),
        stringsAsFactors = FALSE
      )
    }, sanitize.text.function = function(x) x)
  })
  #renders button after a song has been queried
  output$advanced_ui <- renderUI({
    if (knn_done()) {
      actionButton("show_advanced_btn", "Show Advanced KNN Results")
    }
  })
  #re-runs if sogn from dropdown is chosen
  observeEvent(input$alternative_song, {
    if (!selection_made()) {
      selection_made(TRUE)
      return()
    }
    
    req(input$alternative_song)
    selected_song <- input$alternative_song
    selected_match <- top_matches()[top_matches()$song_name == selected_song, ]
    
    best_match(selected_match)
    
    output$current_song_name <- renderText({
      paste("Showing Recommendations for:", best_match()$song_name)
    })
    
    knn_results <- perform_knn(selected_match)
    knn_done(TRUE)
    
    spotify_links <- sapply(knn_results$song_names, get_spotify_link)
    
    output$neighbors <- renderTable({
      data.frame(
        Order = 1:5,
        Song_Name = knn_results$song_names,
        URLs = sprintf('<a href="%s" target="_blank">Link</a>', spotify_links),
        stringsAsFactors = FALSE
      )
    }, sanitize.text.function = function(x) x)
  })
  
  # re-run knnx and observe distance matrix. Fix this...
  observeEvent(input$show_advanced_btn, {
    selected_match <- best_match()
    
    knn_results <- perform_knn(selected_match)
    
    spotify_links <- sapply(knn_results$song_names, get_spotify_link)
    
    output$neighbors <- renderTable({
      data.frame(
        Neighbor_Index = knn_results$indices,
        Distance = knn_results$distances,
        Neighbor_Song = knn_results$song_names,
        URLs = sprintf('<a href="%s" target="_blank">Link</a>', spotify_links),
        stringsAsFactors = FALSE
      )
    }, sanitize.text.function = function(x) x)
  })
  
} 
# Run the app
shinyApp(ui = ui, server = server)




